### 2019-09-20 bp_results_1.tex
updating intro with gail's comments
fixing typos
updated results

\begin{figure}
  \centering
      \includegraphics[width=0.5\textwidth]{dailystress_s1.pdf}
      \includegraphics[width=0.5\textwidth]{dailystress_s3.pdf}
  \caption{The day-to-day stress levels reported by two participants (S1 and S3) are shown. The y-axis represents values on the 5-point Likert scale we asked participants to respond with ranging from 1/Not at all stressed to 5/Extremely stressed. The x-axis represents the day of the study on which the value was recorded (from 0-45). The x-axes are slightly different for the 2 participants as they did not begin and end the study on exactly the same days.}
   \vspace*{-2mm}
\end{figure}

@@

@@

@@

There are many different algorithms that can be used to build a classifier. To select an
algorithm, we compared multiple classifiers using the popular machine learning library scikit-learn~\cite{pedregosa11} and performed a grid search analysis to determine the optimal hyperparameters for each classifier. Our analysis showed that random forest outperforms all other classifiers, including Na\"ive Bayes, decision trees, support vector machine, and a multilayer perceptron neural network. The optimal values for random forest and the three output measures are listed in Table \ref{tab:hyperparams}. For the remainder of this paper, we refer to random forest classifiers trained with these hyperparameters.%\\[-0.1cm]
% for stress they were (minimum samples for split = 4, # of estimators = 100, max features = 0.5, k=300), for focus they were (minimum samples for split = 4, # of estimators = 50, max features = 0.5, k=200), and for awakeness they were (minimum samples for split = 4, # of estimators = 100, max features = 0.25, k=800)
\begin{table}[h]
	\begin{centering}
	\small\addtolength{\tabcolsep}{-1pt}
    \begin{tabular}{llll}
      \hline
      Variable & K & \# Estimators & Minimum Samples Split \\
      \hline
      Stress & 300 & 100 & 4\\
      Focus & 200 & 50 & 4\\
      Awakeness & 800 & 100 & 4\\
      \hline
    \end{tabular}
    \caption{The hyperparameters selected by grid search analysis to tune our random forest models. K refers to the number of features selected.}    \label{tab:hyperparams}
    \end{centering}
\end{table}

\subsection{Individual Classifiers}
%\noindent\textit{Results of Individual Classifiers}\\
Since peoples' experience of stress, focus, and awakeness (as well as
their physiological manifestations) can vary substantially
(e.g.,~\cite{Hernandez11}), we first trained individual classifiers
for each participant (rather than a general one for all
participants). The results of our analysis are reported in
Table~\ref{tab:accuracy}. For our analysis, we report values of
accuracy, one of the most commonly used metric to compare performance,
as well as precision and recall of the classes of interest:
'stressed', 'not focused', and 'not awake'. Since the imbalance in the
data can lead to high accuracy values if a classifier always just
predicts the most likely/frequent class while ignoring the class of
higher importance and interest, precision and recall of the class of
interest are also important to
consider~\cite{yap2014,bhattacharyya_data_2011,Hernandez11}.

@@

Overall, we were able to use extracted physiological features to
predict all three aspects with reasonable accuracy, precision, and
recall, as well as to improve upon the baseline---a stratified random
classifier that randomly chooses one of the two classes with a bias
towards the larger class. While the individually trained classifiers
improved on average across all participants upon the baseline in all
cases excepting recall of 'not focused', the improvement was
substantially higher for awakeness (85\% improvement in precision,
62\% in recall, and 7\% in accuracy) than for stress or focus. Also,
the performance of the individually trained classifiers varied greatly
across participants. While some participants showed a large
improvement, for others the baseline performed much better than the
individually trained classifier. For instance, for predicting
'stressed', the individual classifiers improved upon the baseline for
S4, S6, S8, S11, S12, and S14 with a maximum improvement of 88.4\% in
precision and 184.6\% in recall for S12, while they did worse for S1,
S3, S5, S7, S9, S10, and S13, and in the worst cases did not correctly
predict a single instance of 'stressed'. Typically users that have the
lowest precision and recall values are those where the data is the
most unbalanced. This translates to a scenario where the classifier
will likely label the minority instances as part of the majority class
due to the unbalanced ratio of the data.\\[-0.1cm]
% We attribute these discrepancies to the large differences in response distributions between participants, as well as to the subjectivity of self-reporting.
%
%(improvement: 8\% precision, 1\% recall, 8\% accuracy), the individual performance varied greatly among participants. Some participants showed negligible difference in comparison to the baseline classifier, e.g. subject ..., while others showed a large improvement, e.g. subject ...
%\begin{table}[h!]
%\begin{centering}
%\begin{tabular}{lll}
%\hline
%Participant & Recall Improvement (\%) & Precision Improvement (\%)\\
%\hline
%S12 & 184.6 & 88.4 \\
%S6 & 66.7 & 85.8 \\
%S11 & 50.0 & 44.4 \\
%S8 & 32.4 & 26.7 \\
%S14 & 14.7 & 9.8 \\
%S4 & 5.6 & 5.5 \\
%S9 & -55.4 & -46.8 \\
%S3 & -90.0 & -82.1 \\
%S1 & -100.0 & -100.0 \\
%S5 & -100.0 & -100.0 \\
%S7 & -100.0 & -100.0 \\
%S10 & -100.0 & -100.0 \\
%S13 & -100.0 & -100.0\\
%S2 & - & -\\
%\hline
%\end{tabular}
%\caption{Percentage improvement in recall and precision for stress, using our approach compared to the baseline, on an individual level. For participant 2, the baseline achieved a precision and recall of 0, thus the improvement is undefined}
%\end{centering}
%\label{tab:indImprovement}
%\end{table}

>>>>

\begin{figure}
  \centering
      \includegraphics[width=0.5\textwidth]{dailystress_s1.pdf}
      \includegraphics[width=0.5\textwidth]{dailystress_s3.pdf}
  \caption{The day-to-day stress levels reported by two participants (S1 and S3) are shown. The y-axis represents values on the 5-point Likert scale we asked participants to respond with ranging from 1/Not at all stressed to 5/Extremely stressed. The x-axis represents the day of the study on which the value was recorded (from 0-45). The x-axes are slightly different for the 2 participants as they did not begin and end the study on exactly the same days.}
   \vspace*{-2mm}
   \label{fig:dailyStress}
\end{figure}

@@

Figure \ref{fig:dailyStress} illustrates these points, showing both participants tendency to report and return to baseline stress levels

@@

\begin{table}[h]
	\begin{centering}
	\small\addtolength{\tabcolsep}{-1pt}
    \begin{tabular}{llll}
      \hline
      Variable & K & \# Estimators & Minimum Samples Split \\
      \hline
      Stress & 300 & 100 & 4\\
      Focus & 200 & 50 & 4\\
      Awakeness & 800 & 100 & 4\\
      \hline
    \end{tabular}
    \caption{The hyperparameters selected by grid search analysis to tune our random forest models. K refers to the number of features selected.}    \label{tab:hyperparams}
    \end{centering}
\end{table}

@@

There are many different algorithms that can be used to build a classifier. To select an
algorithm, we compared multiple classifiers using the popular machine learning library scikit-learn~\cite{pedregosa11} and performed a grid search analysis to determine the optimal hyperparameters for each classifier. Our analysis showed that random forest outperforms all other classifiers, including Na\"ive Bayes, decision trees, support vector machine, and a multilayer perceptron neural network. The optimal values for random forest and the three output measures are listed in Table \ref{tab:hyperparams}. For the remainder of this paper, we refer to random forest classifiers trained with these hyperparameters.%\\[-0.1cm]
% for stress they were (minimum samples for split = 4, # of estimators = 100, max features = 0.5, k=300), for focus they were (minimum samples for split = 4, # of estimators = 50, max features = 0.5, k=200), and for awakeness they were (minimum samples for split = 4, # of estimators = 100, max features = 0.25, k=800)

\subsection{Individual Classifiers}
%\noindent\textit{Results of Individual Classifiers}\\
Since peoples' experience of stress, focus, and awakeness (as well as
their physiological manifestations) can vary substantially
(e.g.,~\cite{Hernandez11}), we first trained individual classifiers
for each participant (rather than a general one for all
participants). The results of our analysis are reported in
Table~\ref{tab:accuracy}. For our analysis, we report values of
accuracy, one of the most commonly used metric to compare performance,
as well as precision and recall of the classes of interest:
'stressed', 'not focused', and 'not awake'. Since the imbalance in the
data can lead to high accuracy values if a classifier always just
predicts the most likely/frequent class while ignoring the class of
higher importance and interest, precision and recall of the class of
interest are also important to
consider~\cite{yap2014,bhattacharyya_data_2011,Hernandez11}.
For some users (i.e., S1. as seen In Table \ref{responseDistribution}), the imbalance in their data was so extreme that even after adjusting by oversampling it was impossible to create a reasonable classifier. These scenarios are difficult to predict as any classifier will not have an appropriate balance of stress and not stressed situations to learn from.

@@

Overall, we were able to use extracted physiological features to
predict all three aspects with reasonable accuracy, precision, and
recall, as well as to improve upon the baseline---a stratified random
classifier that randomly chooses one of the two classes with a bias
towards the larger class. While the individually trained classifiers
improved on average across all participants upon the baseline in all
cases excepting recall of 'not focused', the improvement was
substantially higher for awakeness (85\% improvement in precision,
62\% in recall, and 7\% in accuracy) than for stress or focus. Also,
the performance of the individually trained classifiers varied greatly
across participants. While some participants showed a large
improvement, for others the baseline performed much better than the
individually trained classifier. For instance, for predicting
'stressed', the individual classifiers improved upon the baseline for
S4, S6, S8, S11, S12, and S14 with a maximum improvement of 88.4\% in
precision and 184.6\% in recall for S12, while they did worse for S1,
S3, S5, S7, S9, S10, and S13, and in the worst cases did not correctly
predict a single instance of 'stressed'. Typically users that have the
lowest precision and recall values are those where the data is the
most unbalanced.\\[-0.1cm]
% We attribute these discrepancies to the large differences in response distributions between participants, as well as to the subjectivity of self-reporting.
%
%(improvement: 8\% precision, 1\% recall, 8\% accuracy), the individual performance varied greatly among participants. Some participants showed negligible difference in comparison to the baseline classifier, e.g. subject ..., while others showed a large improvement, e.g. subject ...
%\begin{table}[h!]
%\begin{centering}
%\begin{tabular}{lll}
%\hline
%Participant & Recall Improvement (\%) & Precision Improvement (\%)\\
%\hline
%S12 & 184.6 & 88.4 \\
%S6 & 66.7 & 85.8 \\
%S11 & 50.0 & 44.4 \\
%S8 & 32.4 & 26.7 \\
%S14 & 14.7 & 9.8 \\
%S4 & 5.6 & 5.5 \\
%S9 & -55.4 & -46.8 \\
%S3 & -90.0 & -82.1 \\
%S1 & -100.0 & -100.0 \\
%S5 & -100.0 & -100.0 \\
%S7 & -100.0 & -100.0 \\
%S10 & -100.0 & -100.0 \\
%S13 & -100.0 & -100.0\\
%S2 & - & -\\
%\hline
%\end{tabular}
%\caption{Percentage improvement in recall and precision for stress, using our approach compared to the baseline, on an individual level. For participant 2, the baseline achieved a precision and recall of 0, thus the improvement is undefined}
%\end{centering}
%\label{tab:indImprovement}
%\end{table}
